SecFFT: Safeguarding Federated Fine-tuning for Large Vision Language Models against Covert Backdoor Attacks in IoRT Networks

### 2024-09-19 main-v2.tex

\documentclass[lettersize,journal]{IEEEtran}
\usepackage{amsmath,amsfonts}
%\usepackage{algorithmic}
%\usepackage{algorithm}
\usepackage{array}
\usepackage[caption=false,font=scriptsize,labelfont=sf,textfont=rm]{subfig}
\usepackage{textcomp}
\usepackage{stfloats}
\usepackage{url}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{cite}
\hyphenation{op-tical net-works semi-conduc-tor IEEE-Xplore}
% updated with editorial comments 8/9/2021

@@

\usepackage{dsfont}
\usepackage{amssymb}
\usepackage{amsmath,amsthm,amssymb,amsfonts}
\usepackage{ragged2e}
\usepackage{latexsym, amssymb, verbatim, amsmath}
\usepackage{amsmath,bm}
\usepackage{multirow}
\usepackage{mathtools}
\usepackage{subeqnarray}
\usepackage{cases}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathrsfs}
\usepackage{graphicx}
%\usepackage{subfigure}
\usepackage{booktabs}
\usepackage{color}
\usepackage{cite}
\usepackage[]{chapterbib}
\usepackage{enumerate}

@@

\author{Zan Zhou, %~\IEEEmembership{Student Member,~IEEE}, 
 Bo Wang, Tengfei Li, Shujie Yang, Changqiao Xu, Su Yao, 
% <-this % stops a space
\thanks{Z. Zhou, H. Li, S. Huang, S. Yang, P. Guo, and C. Xu are with the State Key Laboratory of Networking and Switching Technology, Beijing University of Posts and Telecommunications, Beijing 100876, P.R. China. E-mail: \{zan.zhou, lihongjing, swhsz, sjyang, guopeng, cqxu\}@bupt.edu.cn. 
	}  % <-this % stops a space
%\thanks{Y. Zhuang is with the Research Institute of China Telecom, Ave Zhongshan, Guangzhou 510630, China. E-mail: 13316094433@chinatelecom.cn.}% <-this % stops a space
% \thanks{L. Zhong is with the Information Engineering College, Capital Normal University, Beijing 100048, China. E-mail: zhonglj@cnu.edu.cn (Corresponding author).}% <-this % stops a space
\thanks{Su Yao is with Tsinghua University, China. E-mail: XXXX@xxx.}% <-this % stops a space
%\thanks{This paper was produced by the IEEE Publication Technology Group. They are in Piscataway, NJ.}% <-this % stops a space
%\thanks{Manuscript received April 19, 2021; revised August 16, 2021.}
}

@@

\begin{abstract}
%攻击者

Fairness-aware federated learning (FFL) plays a crucial role in miti

@@

\begin{IEEEkeywords}
Federated learning, fairness, privacy, integrity, heterogeneous UAV network.
\end{IEEEkeywords}

@@

%随着具身智能技术的迅猛进步和在智慧城市、智能交通、无人机等领域的广泛应用，机器人节点从执行等单一任务的agent，逐步向集合了感知交互和自主决策的智能体进行了演进，这也成为了未来IORT网络的重要发展趋势之一。而视觉大模型作为一种新兴热门的图像感知和物体识别技术，能够成为支撑智能机器人对周围环境进行认识和理解不可或缺的基础。
With the rapid advancements in embodied intelligence technology and its widespread application in multiple areas such as smart cities, intelligent transportation, and unmanned aerial vehicles, robotic nodes have evolved from agents performing singular tasks to intelligent entities that integrate perception, interaction, and autonomous decision-making jointly. This evolution has become one of the critical developmental trends for Internet of Robotic Things (IoRT) networks. Meanwhile, the Large Vision Language Model (LVLM), as an emerging and popular image perception and object recognition technology, has the potential to be an indispensable foundation supporting intelligent robots in understanding and interpreting their surrounding environments.

%视觉大模型极大地增强了机器人的视觉语义理解能力，使其能够对多模态任务进行更加智能化的决策响应，然而在真实应用环境中，仍存在以下问题：1）由于真实世界的多样性和异构性，通用大模型在特定任务上往往无法较好迁移通用知识，需要进行微调才能显著提升识别精度（例如，针对动目标检测任务，交通机器人主要针对车辆、行人，而工厂机器人则主要针对机床，两者轮廓、行为模式范围均存在较大差异）；2）此外，将来自不同源头的离散化、碎片化分布的海量数据进行传输并汇聚于主服务器，将会导致巨大的通信开销和时间延迟，无法满足具身智能业务的需求，且存在潜在的隐私风险。
Although LVLMs significantly enhance the visual semantic understanding capabilities of robots, enabling them to make more intelligent decisions in multi-modal tasks, there are still challenges in real-world applications: 1) Due to the diversity and heterogeneity of the real world, general foundation models often cannot effectively transfer their general knowledge to specific tasks, necessitating fine-tuning to significantly improve recognition accuracy (for example, traffic robots primarily focus on vehicles and pedestrians for moving object detection tasks, while factory robots mainly target machine tools, with significant differences in their contours and behavioral patterns); 2) Additionally, transmitting and aggregating massive amounts of discretized and fragmented data from different sources to a central server can result in significant communication overhead and latency, failing to meet the demands of embodied intelligence applications and posing potential privacy risks.

@@

%不幸的是，FFT的分布式特性为多种后门攻击提供了存续的空间。尽管多种防御手段相继提出，但是尽我们所知，目前尚未有针对LVM的FFT过程进行污染防治的后门对抗技术提出。现有技术大多是针对普通联邦学习场景或集中式的VLM训练进行设计，这暴露出两大核心问题：（1）高隐蔽的攻击行为感知困难：攻击者在设计攻击行为时，(2)复杂攻击策略的检测困难：

% 因此，在本文中，我们提出了一种
% 主要包括以下几方面贡献：
% 我们首先提出了针对进行防御
% 针对意图
% 防御策略

The remainder of this paper is organized as follows: Section \uppercase\expandafter{\romannumeral2} reviews the related works. Section \uppercase\expandafter{\romannumeral3} and \uppercase\expandafter{\romannumeral4} present the system and threat models, respectively. Section \uppercase\expandafter{\romannumeral6} elaborates on the design philosophy of the proposed SecFFT solution, providing comprehensive details on the availability, security, and integrity of FFT in IoRT networks. Section \uppercase\expandafter{\romannumeral7} presents the experimental results. Finally, Section \uppercase\expandafter{\romannumeral8} concludes this paper and briefly explores future directions.

@@

\begin{figure}[!t]
    \centering
    \includegraphics[width=\linewidth]{figures/SecFFT-scenario.pdf}
    \caption{FFT process for large vision language models in IoRT network}
    \label{fig1:scene}
\end{figure}

@@

% 空间隐蔽性后门攻击是指攻击者通过微调模型参数、修改输入特征或操纵模型训练过程等方式，使得现有防御方法无法在特征空间中有效区分恶意攻击者与良性参与者，从而有效逃避检测。EDGE CASE\cite{EDGE_CASE}提出了一种针对联邦学习的边缘案例后门攻击方法，通过在训练和测试数据集中通常不存在但自然存在的输入数据点（即边缘样本）来注入后门。Shadowcast\cite{xu2024shadowcast}通过在特征空间中引入微小、不可察觉的扰动，操纵视觉语言模型（VLMs）在特定视觉输入下生成误导性的文本输出。Flip\cite{jha2023label}通过修改训练数据的标签，使得模型训练轨迹接近于带有后门的专家模型，从而在特征空间中达到隐蔽攻击的效果，即使在图像干净的情况下也能逃避现有检测方法。
\textbf{Spatially Covert Backdoor Attacks} refer to attacks where adversaries fine-tune model parameters, modify input features, or manipulate the model training process in a way that existing defense cannot effectively distinguish between malicious attackers and benign participants in the feature space, thereby evading detection. EDGE CASE\cite{EDGE_CASE} proposes an edge-case backdoor attack method for federated learning by injecting backdoors using input data points that are naturally occurring but typically absent in both training and testing datasets (i.e., edge-case samples). Shadowcast \cite{xu2024shadowcast} manipulates vision-language models (VLMs) to produce misleading textual outputs for specific visual inputs by introducing subtle, imperceptible perturbations in the feature space. Flip \cite{jha2023label} modifies the labels of training data so that the model's training trajectory approximates that of an expert model with a backdoor, achieving covert attacks in the feature space that can evade existing detection methods even when the images are clean.

% 时间隐蔽性后门攻击是指攻击者通过调整模型训练过程或优化后门植入等方式，使后门在恶意攻击者结束攻击后仍能长期存在，不被后续良性更新所覆盖，从而持续影响模型行为。 Neurotoxin\cite{zhang2022neurotoxin}提出了一种针对联邦学习的更持久的后门攻击方法，通过攻击在训练过程中变化较小的模型参数来增加后门的耐久性，避免与良性用户梯度的冲突，使其不易在后续的模型更新中被清除。ImgTrojan\cite{tao2024imgtrojan}通过在视觉语言模型的训练数据中注入少量恶意样本，使得这些后门即使在模型后续的良性训练中仍能保持有效，从而持续影响模型的行为。\cite{gu2023gradient}提出了一种新的更新控制方法，通过跨层更新幅度归一化和层内更新方向投影，解决了参数高效调优（PET）过程中后门攻击的遗忘问题，从而维持攻击的持久性和隐蔽性。A3FL\cite{zhang2024a3fl}通过对抗性适应损失函数来优化触发器的方式，使得后门能够在全局训练动态中持久存在，从而在联邦学习模型中实现高效且持久的攻击效果。
% Chameleon \cite{dai2023chameleon} focuses on enhancing the spatial covert nature of backdoor attacks by adjusting the relationships between poisoned samples, interference samples, and promoting samples in contrastive learning.
\textbf{Temporally Covert Backdoor Attacks} involve adjusting the model training process or optimizing backdoor implantation such that the backdoor persists even after the malicious attacker has stopped the attack, remaining undetected by subsequent benign updates, and continuing to affect the model’s behavior. NEUR\cite{zhang2022neurotoxin} proposes a more persistent backdoor attack method for federated learning by targeting model parameters that change less during training. This approach increases the durability of the backdoor, avoids conflicts with the gradients of benign users, and makes it less likely to be removed in subsequent model updates. ImgTrojan \cite{tao2024imgtrojan} injects a small number of malicious samples into the training data of vision-language models, ensuring that these backdoors remain effective even after subsequent benign training, thereby continuously affecting the model's behavior. \cite{gu2023gradient} proposes a novel update control method that normalizes cross-layer update magnitudes and projects intra-layer update directions to address the issue of backdoor forgetting during parameter-efficient tuning (PET), thus maintaining the persistence and concealment of the attack. A3FL \cite{zhang2024a3fl} optimizes the trigger by using an adversarial adaptation loss function, enabling the backdoor to persist throughout global training dynamics, thereby achieving efficient and persistent attack effects.
% 总之，现有的后门攻击策略在空间隐蔽性和时间隐蔽性方面都有了显著的进展。空间隐蔽性后门攻击通过优化特征空间中的触发器或更新，使得恶意行为难以被现有的检测方法发现，从而实现了更高效、更难以检测的攻击。而时间隐蔽性后门攻击则专注于提高后门的持久性，使得即使在恶意攻击结束后，这些后门依然能够在模型中长期存在并继续影响其行为。通过这些攻击方式的结合，攻击者能够在各种应用场景中有效地躲避检测，并确保攻击效果的持续性，这对现有的防御机制提出了更高的挑战，强调了未来在开发更加鲁棒的防御方法方面的迫切需求。

%总的来看，现有的后门攻击方法已展现出显著的隐蔽性和持久性，能够有效规避当前的检测机制。这些攻击不仅在特征空间中巧妙隐藏恶意意图，还能在长期训练过程中保持攻击效果，对模型行为产生持续而深远的影响，也对模型安全构成了严峻挑战。这对现有的防御机制提出了更高的挑战，强调了未来在开发更加鲁棒的防御方法方面的迫切需求。
Overall, existing backdoor attack methods have demonstrated significant advances in both spatial and temporal concealment, effectively circumventing current detection mechanisms. These attacks not only cleverly hide malicious intent in the feature space but also maintain their effects throughout long-term training, posing a sustained and profound impact on model behavior, thus presenting a severe challenge to model security. This underscores the need for future research to focus on developing more robust defense methods.

\subsection{Backdoor countermeasures}
%在当前深度学习模型的安全性研究中，针对后门攻击的防御措施主要可以分为两大类：后门检测和后门缓解。% Backdoor Mitigation or Elimination
In the current research on the defense technologies, defense measures against backdoor attacks can be also broadly categorized into two major types: %backdoor detection and backdoor mitigation.

@@

%\textbf{本地模型训练}：
\textbf{Local training}:
    %在每轮联邦训练的开始时，中心服务器将全局模型的增量参数 $\theta_g^{t-1}$ 发送给每个机器人节点$R_i$。每个节点在获得增量参数$\theta_g^{t-1}$后，基于本地数据集 $D_i$ 进行微调训练。
    At the beginning of each federated training round, the central server sends the global model's incremental parameters $\theta_g^{t-1}$ to each robotic node $R_i$. Upon receiving $\theta_g^{t-1}$, each node fine-tunes the model based on its local dataset $D_i$.
    % 算法\ref{alg:local_update}展示了节点进行本地模型每轮训练的主要过程。
    %假设本地训练的轮次为$E_l$，批次大小为$b$，学习率为$\eta$。对于每一轮，节点$R_i$首先从其本地数据集$D_i$中随机抽取一个批次数据$D_i^b \sim Sample(D_i, b)$，然后利用随机梯度下降等优化算法，对增量参数进行更新，
    Assume the number of local training epochs is $E_l$, the batch size is $b$, and the learning rate is $\eta$. For each epoch, the node $R_i$ randomly samples a batch of data $D_i^b \sim Sample(D_i, b)$ from its local dataset and updates the incremental parameters using an optimization algorithm such as stochastic gradient descent (SGD):
\begin{equation}
    (\theta_i^{t})^e = (\theta_i^{t})^{e-1} - \eta \nabla f(D_i^b, (\theta_i^{t})^{e-1}),
\end{equation}
    %其中$(\theta_i^{t})^e$表示在第$e$轮微调后的参数，初始时$(\theta_i^{t})^0 = \theta_g^{t-1}$。
    where $(\theta_i^{t})^e$ denotes the parameters after the $e$-th epoch, with the initial value $(\theta_i^{t})^0 = \theta_g^{t-1}$.

>>>>

\documentclass[lettersize,journal]{IEEEtran}
\usepackage{amsmath,amsfonts}
%\usepackage{algorithmic}
%\usepackage{algorithm}
\usepackage{array}
\usepackage[caption=false,font=scriptsize,labelfont=sf,textfont=rm]{subfig}
\usepackage{textcomp}
\usepackage{stfloats}
\usepackage{url}
\usepackage{verbatim}
\usepackage{graphicx}
% \usepackage{tabularx}
\usepackage{cite}
\hyphenation{op-tical net-works semi-conduc-tor IEEE-Xplore}
% updated with editorial comments 8/9/2021

@@

\usepackage{dsfont}
\usepackage{amssymb}
\usepackage{amsmath,amsthm,amssymb,amsfonts}
\usepackage{ragged2e}
\usepackage{latexsym, amssymb, verbatim, amsmath}
\usepackage{amsmath,bm}
\usepackage{multirow}
\usepackage{mathtools}
\usepackage{subeqnarray}
\usepackage{cases}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathrsfs}
\usepackage{graphicx}
%\usepackage{subfigure}
\usepackage{booktabs}
\usepackage{color}
\usepackage{cite}
\usepackage[]{chapterbib}
\usepackage{enumerate}
\usepackage{threeparttable}

@@

\author{Zan Zhou, %~\IEEEmembership{Student Member,~IEEE}, 
  Changqiao Xu, Bo Wang, Tengfei Li, Sizhe Huang, Shujie Yang$^{\ast}$\thanks{*Corresponding authors}, Su Yao$^{\ast}$
% <-this % stops a space
\thanks{Z. Zhou, B. Wang, T. Li, S. Huang, C. Xu, and S. Yang are with the State Key Laboratory of Networking and Switching Technology, Beijing University of Posts and Telecommunications, Beijing 100876, P.R. China. E-mail: \{zan.zhou, cqxu, bowang, tf, swhsz, sjyang\}@bupt.edu.cn. 
	}  % <-this % stops a space
%\thanks{Y. Zhuang is with the Research Institute of China Telecom, Ave Zhongshan, Guangzhou 510630, China. E-mail: 13316094433@chinatelecom.cn.}% <-this % stops a space
% \thanks{L. Zhong is with the Information Engineering College, Capital Normal University, Beijing 100048, China. E-mail: zhonglj@cnu.edu.cn (Corresponding author).}% <-this % stops a space
\thanks{Su Yao is with the Beijing National Research Center for Information Science and Technology (BNRist), Tsinghua University, Beijing 100190, China. E-mail: yaosu@tsinghua.edu.cn}% <-this % stops a space
%\thanks{This paper was produced by the IEEE Publication Technology Group. They are in Piscataway, NJ.}% <-this % stops a space
%\thanks{Manuscript received April 19, 2021; revised August 16, 2021.}
}

@@

\begin{abstract}
%随着大模型与具身智能机器人网络的迅速发展，以及在智慧城市、电网、工厂、交通运输等领域的应用，视觉感知理解成为了突破具身智能节点性能瓶颈的基础性要素之一。而由于通用的预训练模型并不能很好应对特定任务，联邦微调技术由于其能够最大限度利用离散节点数据、算力的优势，成为了提升视觉感知模型能力的热带技术。然而，随着多种针对性高级持续性威胁的涌现，现有的污染防治方案，并不能有效应对这一新场景下的隐蔽后门攻击。本文因此提出SecFFT方法，其核心在于针对攻击隐蔽性高与攻击策略复杂的问题，分别设计了基于频域分布一致性提取的瞬时攻击行为识别，和基于隐藏攻击意图检测的长效安全聚合机制。从最终目的出发，有效限缩了智能化攻击者通过隐藏自身行为来绕过防御方案的可行域。在公共数据集上的实验表明，我们所提出的方法能够显著提升面对后门攻击时的防御成功率、模型性能和识别精度，尤其是面对具有多轮次复杂策略的高隐蔽后门范式。
As the large vision language models and embodied intelligent robotic networks continue to advance at a remarkable pace, particularly in applications spanning smart cities, power grids, factories, and transportation, visual perception and understanding have emerged as foundational elements to overcoming performance limitations in such intelligent systems. However, since general pre-trained models are not well-suited to specific tasks, federated fine-tuning (FFT) has gained attention as a promising technique for enhancing the performance of vision-based perception models by leveraging data and computational power distributed across nodes. The rise of advanced persistent threats has revealed significant vulnerabilities in existing defense mechanisms, which struggle to mitigate sophisticated backdoor attacks toward FFT for LVLMs. To address these challenges, this paper proposes the SecFFT method, which tackles both the stealthiness and complexity of backdoor strategies. The approach incorporates instantaneous attack behavior detection based on frequency-domain distribution consistency and introduces a long-term secure aggregation mechanism aimed at identifying hidden attack intentions. These strategies effectively limit the feasibility of adversaries attempting to bypass defense measures by concealing their behaviors. Experiments conducted on public datasets demonstrate that SecFFT significantly improves defense success rates, model performance, and detection accuracy, particularly in response to highly covert, multi-round backdoor attacks.

@@

\begin{IEEEkeywords}
Vision Language Model, Federated learning, Fine-tuning, Security, Backdoor, Internet of Robotic Things.
\end{IEEEkeywords}

@@

%随着具身智能技术的迅猛进步和在智慧城市、智能交通、无人机等领域的广泛应用，机器人节点从执行等单一任务的agent，逐步向集合了感知交互和自主决策的智能体进行了演进，这也成为了未来IORT网络的重要发展趋势之一。而视觉大模型作为一种新兴热门的图像感知和物体识别技术，能够成为支撑智能机器人对周围环境进行认识和理解不可或缺的基础。
With the rapid advancements of embodied intelligence technology and its widespread application in multiple areas such as smart cities, intelligent transportation, and unmanned aerial vehicles, robotic nodes have evolved from agents performing singular tasks to intelligent entities that integrate perception, interaction, and autonomous decision-making. This evolution has become one of the critical developmental trends for Internet of Robotic Things (IoRT) networks\cite{andronie2023big}. Meanwhile, the Large Vision Language Model (LVLM)\cite{liu2024survey}, as an emerging and popular image perception and object recognition technology, has the potential to support intelligent robots in understanding and interpreting their surrounding environments.

%视觉大模型极大地增强了机器人的视觉语义理解能力，使其能够对多模态任务进行更加智能化的决策响应，然而在真实应用环境中，仍存在以下问题：1）由于真实世界的多样性和异构性，通用大模型在特定任务上往往无法较好迁移通用知识，需要进行微调才能显著提升识别精度（例如，针对动目标检测任务，交通机器人主要针对车辆、行人，而工厂机器人则主要针对机床，两者轮廓、行为模式范围均存在较大差异）；2）此外，将来自不同源头的离散化、碎片化分布的海量数据进行传输并汇聚于主服务器，将会导致巨大的通信开销和时间延迟，无法满足具身智能业务的需求，且存在潜在的隐私风险。
Although LVLMs significantly enhance the visual semantic understanding capabilities of robots, enabling them to make more intelligent decisions in multi-modal tasks, there are still challenges in real-world applications: (1) Due to the diversity and heterogeneity of the real world\cite{dai2023tackling}, general foundation models often cannot effectively transfer their general knowledge to specific tasks, necessitating fine-tuning\cite{fu2023effectiveness} to significantly improve recognition accuracy (for example, traffic robots primarily focus on vehicles and pedestrians for moving object detection tasks, while factory robots mainly target machine tools, with significant differences in their contours and behavioral patterns); (2) Additionally, transmitting and aggregating massive amounts of discretized and fragmented data from different sources to a central server can result in significant communication overhead and latency\cite{zhao2023towards}, failing to meet the demands of embodied intelligence applications and posing potential privacy risks\cite{wen2023survey, khan2023federated}.

@@

%不幸的是，FFT的分布式特性为多种后门攻击提供了存续的空间。尽管多种防御手段相继提出，但是尽我们所知，目前尚未有针对LVM的FFT过程进行污染防治的后门对抗技术提出。现有技术大多是针对普通联邦学习场景或集中式的VLM训练进行设计，这暴露出两大核心问题：（1）高隐蔽的攻击行为感知困难：攻击者在设计攻击行为时，尽量在行为特征上拟合正常用户，perception的区分度越来越narrow；(2)复杂攻击策略的检测困难：为了规避防御，攻击者可能采用限制角度、大小、符号等方式，将原有单次攻击转化为多轮次长时间尺度的持续性威胁，攻击识别更加困难。
Unfortunately, the distributed nature of FFT provides a persistent space for various backdoor attacks. While several defense strategies have been proposed, to the best of our knowledge, no countermeasures specifically address backdoor attacks in the FFT process for LLVMs. Most existing techniques are designed for conventional federated learning scenarios or centralized VLM training, exposing two critical issues: (1) \textbf{The high stealthy nature of attack behavior:} Attackers increasingly design their actions to mimic normal users, narrowing the perception’s distinguishing capability\cite{xu2024shadowcast, zhang2024a3fl}; and (2) \textbf{The complexity of detecting sophisticated attack strategies:} To evade defenses, attackers may employ techniques such as limiting angle, magnitude, or sign, transforming a single attack into a multi-round, long-term threat, further complicating attack detection\cite{geisler2024attacking, chen2024optimal, sagliano2024powered, dong2023adaptive,wan2023average}.

% 因此，在本文中，我们针对LVLM的FFT过程，提出了一种新型的抗backdoor防御方案。针对攻击行为感知难题，我们创新地从频域分布一致性入手，将隐性的梯度更新分布差异而非显性的更新离群性作为新的鉴别角度，从而实现对高隐蔽攻击的感知；在此基础上，进一步针对智能化攻击者可能采取的复杂多轮次策略，我们创新地提出了基于意图识别的长效评估架构，有效分离高级持续性威胁，并据此设计安全聚合方法。
% 主要包括以下几方面贡献：
Therefore, in this paper, we propose a novel backdoor defense scheme tailored for the FFT process of LVLMs. To address the challenge of detecting attack behaviors, we introduce a groundbreaking approach based on consistency in frequency domain distributions. Rather than focusing on explicit gradient outliers, we leverage the implicit differences in gradient update distributions as a new identification criterion, enabling the detection of highly covert attacks. Building on this, we further address the issue of complex multi-round strategies employed by intelligent attackers by proposing an innovative long-term evaluation framework based on intent recognition. This framework effectively isolates advanced persistent threats and informs the design of a secure aggregation method.

The main contributions of this scheme include:
\begin{enumerate}
    \item Based on an in-depth analysis of emerging covert backdoor attacks, we first define two core challenges: detecting highly covert attack behaviors and identifying complex attack strategies. We then provide a case study of these issues within the FFT process of large vision models. To the best of our knowledge, this is the first framework capable of recognizing deep-level attack intentions during the fine-tuning process in IoRT networks, effectively defending against covert attacks.
    \item In response to the potential threat of intelligent attackers fragmenting their behaviors to evade detection, we propose an innovative method for constructing attack intentions based on historical behavior sequences. By building an approximate minimal covering hypersphere from these behaviors, the malicious intent hidden behind seemingly disordered, benign actions is fully exposed. 
    \item SecFFT also adopts a novel nonlinear defense strategy, dynamically adjusting aggregation weights using local outlier factors while ensuring fairness for benign nodes and effectively eliminating malicious ones. A series of experiments on public datasets demonstrate that our approach significantly improves model integrity, effectively suppresses covert backdoor attacks, and maintains a low false positive rate.
\end{enumerate}
% 基于对现有新型隐蔽后门攻击的分析，我们首先定义了高隐蔽攻击行为感知和复杂攻击策略识别这两大核心问题，并基于视觉大模型的FFT过程进行实例分析。据我们所知，这是首个在IORT网络中能够实现对微调过程中深层次攻击目的识别，从而抵御高隐蔽攻击的方案。
% 针对智能攻击者可能将攻击行为切片化拆解从而规避防御检测的潜在威胁，我们创新地设计了基于历史行为序列进行攻击意图构建的方法。通过构建历史行为的近似最小覆盖超球，将智能攻击者隐藏在看似杂乱无序、良性无害的攻击行为背后的恶意意图充分暴露。
% 基于意图识别架构，SecFFT能够设计一种非线性的新型防御策略，利用局部离群因子动态调整聚合权重，同时保障良性用户公平性和恶意用户的消除。公有数据集上的系列实验表明，我们的方法能够显著提升模型完整性，有效遏制隐蔽后门攻击，并同时保证极低的误判率。

The remainder of this paper is organized as follows: Section \uppercase\expandafter{\romannumeral2} reviews related works. Section \uppercase\expandafter{\romannumeral3} and \uppercase\expandafter{\romannumeral4} present the system and threat models, respectively. Section \uppercase\expandafter{\romannumeral6} elaborates on the design philosophy of the proposed SecFFT solution, providing comprehensive details on the availability, security, and integrity of FFT in IoRT networks. Section \uppercase\expandafter{\romannumeral7} presents the experimental results. Finally, Section \uppercase\expandafter{\romannumeral8} concludes this paper and briefly explores future directions.

@@

\begin{figure}[!t]
    \centering
    \includegraphics[width=\linewidth]{figures/SecFFT-scenario.pdf}
    \caption{FFT process for large vision language models in IoRT network}
    \label{fig1:scene}
    %\vspace{-0.7cm}
\end{figure}

@@

% 空间隐蔽性后门攻击是指攻击者通过微调模型参数、修改输入特征或操纵模型训练过程等方式，使得现有防御方法无法在特征空间中有效区分恶意攻击者与良性参与者，从而有效逃避检测。EDGE CASE\cite{EDGE_CASE}提出了一种针对联邦学习的边缘案例后门攻击方法，通过在训练和测试数据集中通常不存在但自然存在的输入数据点（即边缘样本）来注入后门。Shadowcast\cite{xu2024shadowcast}通过在特征空间中引入微小、不可察觉的扰动，操纵视觉语言模型（VLMs）在特定视觉输入下生成误导性的文本输出。Flip\cite{jha2023label}通过修改训练数据的标签，使得模型训练轨迹接近于带有后门的专家模型，从而在特征空间中达到隐蔽攻击的效果，即使在图像干净的情况下也能逃避现有检测方法。
\textbf{Spatially Covert Backdoor Attacks} refer to attacks where adversaries fine-tune model parameters, modify input features, or manipulate the model training process in a way that existing defense cannot effectively distinguish between malicious attackers and benign participants in the feature space, thereby evading detection. EDGE CASE\cite{EDGE_CASE} proposes an edge-case backdoor attack method for federated learning by injecting backdoors using input data points that are naturally occurring but typically absent in both training and testing datasets (i.e., edge-case samples). Shadowcast \cite{xu2024shadowcast} manipulates VLMs to produce misleading textual outputs for specific visual inputs by introducing subtle, imperceptible perturbations in the feature space. Flip \cite{jha2023label} modifies the labels of training data so that the model's training trajectory approximates that of an expert model with a backdoor. This hidden process completes covert attacks in the feature space, thus evading existing detection methods even when the images are clean.

% 时间隐蔽性后门攻击是指攻击者通过调整模型训练过程或优化后门植入等方式，使后门在恶意攻击者结束攻击后仍能长期存在，不被后续良性更新所覆盖，从而持续影响模型行为。 Neurotoxin\cite{zhang2022neurotoxin}提出了一种针对联邦学习的更持久的后门攻击方法，通过攻击在训练过程中变化较小的模型参数来增加后门的耐久性，避免与良性用户梯度的冲突，使其不易在后续的模型更新中被清除。ImgTrojan\cite{tao2024imgtrojan}通过在视觉语言模型的训练数据中注入少量恶意样本，使得这些后门即使在模型后续的良性训练中仍能保持有效，从而持续影响模型的行为。\cite{gu2023gradient}提出了一种新的更新控制方法，通过跨层更新幅度归一化和层内更新方向投影，解决了参数高效调优（PET）过程中后门攻击的遗忘问题，从而维持攻击的持久性和隐蔽性。A3FL\cite{zhang2024a3fl}通过对抗性适应损失函数来优化触发器的方式，使得后门能够在全局训练动态中持久存在，从而在联邦学习模型中实现高效且持久的攻击效果。
% Chameleon \cite{dai2023chameleon} focuses on enhancing the spatial covert nature of backdoor attacks by adjusting the relationships between poisoned samples, interference samples, and promoting samples in contrastive learning.
\textbf{Temporally Covert Backdoor Attacks} involve adjusting the model training process or optimizing backdoor implantation such that the backdoor persists even after the malicious attacker has stopped the attack, remaining undetected by subsequent benign updates, and continuing to affect the model’s behavior. NEUR\cite{zhang2022neurotoxin} proposes a more persistent backdoor attack method for federated learning by targeting model parameters that change less during training. This approach increases the durability of the backdoor, avoids conflicts with the gradients of benign nodes, and makes it less likely to be removed in subsequent model updates. ImgTrojan \cite{tao2024imgtrojan} injects a small number of malicious samples into the training data of vision-language models, ensuring that these backdoors remain effective even after subsequent benign training, thereby continuously affecting the model's behavior. \cite{gu2023gradient} proposes a novel update control method that normalizes cross-layer update magnitudes and projects intra-layer update directions to address the issue of backdoor forgetting during parameter-efficient tuning (PET), thus maintaining the persistence and concealment of the attack. A3FL \cite{zhang2024a3fl} optimizes the trigger by using an adversarial adaptation loss function, enabling the backdoor to persist throughout global training dynamics, thereby achieving efficient and persistent attack effects.
% 总之，现有的后门攻击策略在空间隐蔽性和时间隐蔽性方面都有了显著的进展。空间隐蔽性后门攻击通过优化特征空间中的触发器或更新，使得恶意行为难以被现有的检测方法发现，从而实现了更高效、更难以检测的攻击。而时间隐蔽性后门攻击则专注于提高后门的持久性，使得即使在恶意攻击结束后，这些后门依然能够在模型中长期存在并继续影响其行为。通过这些攻击方式的结合，攻击者能够在各种应用场景中有效地躲避检测，并确保攻击效果的持续性，这对现有的防御机制提出了更高的挑战，强调了未来在开发更加鲁棒的防御方法方面的迫切需求。

%总的来看，现有的后门攻击方法已展现出显著的隐蔽性和持久性，能够有效规避当前的检测机制。这些攻击不仅在特征空间中巧妙隐藏恶意意图，还能在长期训练过程中保持攻击效果，对模型行为产生持续而深远的影响，也对模型安全构成了严峻挑战。这对现有的防御机制提出了更高的挑战，强调了未来在开发更加鲁棒的防御方法方面的迫切需求。
Overall, existing backdoor attack methods have demonstrated significant advances in both spatial and temporal concealment, effectively circumventing current detection mechanisms. These attacks not only cleverly hide malicious intent in the feature space but also maintain their effects throughout long-term training, posing a sustained and profound impact on model behavior, thus presenting a severe challenge to model security. This underscores the need for developing more robust defense methods.

\subsection{Backdoor countermeasures}
%在当前深度学习模型的安全性研究中，针对后门攻击的防御措施主要可以分为两大类：后门检测和后门缓解。% Backdoor Mitigation or Elimination
In current research on defense technologies, defense measures against backdoor attacks can be also broadly categorized into two major types: %backdoor detection and backdoor mitigation.

@@

%\textbf{本地模型训练}：
\textbf{Local training}:
    %在每轮联邦训练的开始时，中心服务器将全局模型的增量参数 $\theta_g^{t-1}$ 发送给每个机器人节点$R_i$。每个节点在获得增量参数$\theta_g^{t-1}$后，基于本地数据集 $D_i$ 进行微调训练。
    At the beginning of each federated training round, the central server sends the global model's incremental parameters $\theta_g^{t-1}$ to each robotic node $R_i$. Upon receiving $\theta_g^{t-1}$, each node fine-tunes the model based on its local dataset $D_i$.
    % 算法\ref{alg:local_update}展示了节点进行本地模型每轮训练的主要过程。
    %假设本地训练的轮次为$E_l$，批次大小为$b$，学习率为$\eta$。对于每一轮，节点$R_i$首先从其本地数据集$D_i$中随机抽取一个批次数据$D_i^b \sim Sample(D_i, b)$，然后利用随机梯度下降等优化算法，对增量参数进行更新，
    Assume the number of local training epochs is $E_l$, the batch size is $b$, and the learning rate is $\eta$. For each epoch, the node $R_i$ randomly samples a batch of data $D_i^b \sim Sample(D_i, b)$ from its local dataset and updates the incremental parameters using an optimization procedure as:
\begin{equation}
    (\theta_i^{t})^e = (\theta_i^{t})^{e-1} - \eta \nabla f(D_i^b, (\theta_i^{t})^{e-1}),
\end{equation}
    %其中$(\theta_i^{t})^e$表示在第$e$轮微调后的参数，初始时$(\theta_i^{t})^0 = \theta_g^{t-1}$。
    where $(\theta_i^{t})^e$ denotes the parameters after the $e$-th epoch, with the initial value $(\theta_i^{t})^0 = \theta_g^{t-1}$.

    %本地训练的目标是最小化损失函数 $f(D_i^b, (\theta_i^{t})^e)$：
    The local training objective to minimize the loss $f(D_i^b, (\theta_i^{t})^e)$ is:
\begin{equation}
    f(D_i^b, (\theta_i^{t})^e) = \frac{1}{|D_i^b|} \sum_{j=1}^{|D_i^b|} \ell(x_{i,j}, y_{i,j}; \Theta_p + (\theta_i^{t})^e),
\end{equation}
    %其中$\Theta$是模型的原始参数， $\ell$表示损失函数（如交叉熵损失）。本地训练完成后，节点计算其增量参数更新 $\triangledown \theta_i^t = (\theta_i^t)^{E_l} - (\theta_i^t)^0$ 并上传至中心服务器。
    where $\Theta_p$ represents the original parameters of the model, and $\ell$ denotes the loss function (e.g., cross-entropy loss). After local training is completed, the node computes the update to the incremental parameters $\triangledown \theta_i^t = (\theta_i^t)^{E_l} - (\theta_i^t)^0$ and uploads the result to the central server.
